---
title: "StatInf_Project1_Simulation_Exercise"
author: "Willianto Asalim"
date: "21/05/2020"
output:
  html_document: default
  pdf_document: default
---



##  Statistical Inference Proj 1: Simulation Exercise
***

Investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations.  Investigate the distribution of averages of 40 exponentials and perform a thousand simulations.

To illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials by performing the following tasks.
- Show the sample mean and compare it to the theoretical mean of the distribution.
- Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
- Show that the distribution is approximately normal.


```{r LoadPackages, echo=FALSE}
library(knitr) ##Load Knitr package
library(ggplot2) ##Plotting and data
library(pander) ##Load package for tidy T test result
```

```{r setoptions, echo=FALSE}
## Setting Global Option where echo = true so that someone will be able to read the code and results.
knitr::opts_chunk$set(echo = TRUE, results = "hold", tidy = TRUE)
```
<br />

#### 1. Show the sample mean and compare it to the theoretical mean of the distribution.
***


#### 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
***


#### 3. Show that the distribution is approximately normal.
***


#### 4. Appendix
***
The platform specification used:

Spec    | Description
------- | -----------------------
OS      | Windows 10 Pro - 64 bit
CPU     | AMD Ryzen 5 - 3400G
RAM     | 16GB DDR4 3000MHz
Storage | 500GB SSD - M.2 NVMe (PCIe) 

