---
title: "StatInf_Project1_Simulation_Exercise"
author: "Willianto Asalim"
date: "21/05/2020"
output:
  html_document: default
  pdf_document: default
    keep_md: yes
---



##  Statistical Inference Proj 1: Simulation Exercise
***

Investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations.  Investigate the distribution of averages of 40 exponentials and perform a thousand simulations.

To illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials using the following:
- Show the sample mean and compare it to the theoretical mean of the distribution.
- Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
- Show that the distribution is approximately normal.


```{r LoadPackages, echo=FALSE}
library(knitr) ##Load Knitr package
library(ggplot2) ##Plotting and data
library(pander) ##Load package for tidy T test result
```

```{r setoptions, echo=FALSE}
## Setting Global Option where echo = true so that someone will be able to read the code and results.
knitr::opts_chunk$set(echo = TRUE, results = "hold", tidy = TRUE)
```
<br />


```{r SimulationSetup}
##Setup variables provided by the instructions.
n <- 40 ##number of exponentials
lambda <- 0.2 ##universal lambda for all simulations
simNum <- 1000 ## number of simulations

set.seed(1234) ## set the seed to create reproducibility

##Simulation in matrix format
sim <- matrix(rexp(simNum * n, rate=lambda), simNum, n)
meanSim <- rowMeans(sim) ## Refer to Appendix
```


#### 1. Show the sample mean and compare it to the theoretical mean of the distribution.
***
As stated in the instruction the expected mean, mu of a exponential distribution is 1/lambda.We are going to compare this with the distribution mean.

```{r Theoretical mean}
muTheory <- 1/lambda
muTheory
```


```{r SampleMean}
muSim <-mean(meanSim)
muSim
```

```{r SampleMeanHistogram}
hist(meanSim, xlab = "Mean", main = "Sample Mean Distribution Simulation")
abline(v = muSim, col = "red")
abline(v = muTheory, col = "green")
```

The sample mean is `r muSim` and the theoretical mean is `r muTheory` which show that they are very close. This can be further seen with the histogram above.

#### 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
***


#### 3. Show that the distribution is approximately normal.
***


#### 4. Appendix
***

First 6 lines of the mean of simulation
`r head(meanSim)`

The platform specification used:

Spec    | Description
------- | -----------------------
OS      | Windows 10 Pro - 64 bit
CPU     | AMD Ryzen 5 - 3400G
RAM     | 16GB DDR4 3000MHz
Storage | 500GB SSD - M.2 NVMe (PCIe) 

