---
title: "StatInf_Project1_Simulation_Exercise"
author: "Willianto Asalim"
date: "21/05/2020"
output:
  pdf_document: default
  html_document: default
---



##  Statistical Inference Proj 1: Simulation Exercise
***

Investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations.  Investigate the distribution of averages of 40 exponentials and perform a thousand simulations.

To illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials using the following:
<br />

- Show the sample mean and compare it to the theoretical mean of the distribution.
- Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
- Show that the distribution is approximately normal.

<br />

```{r LoadPackages, echo=FALSE}
library(knitr) ##Load Knitr package
library(ggplot2) ##Plotting and data
library(pander) ##Load package for tidy T test result
```

```{r setoptions, echo=FALSE}
## Setting Global Option where echo = true so that someone will be able to read the code and results.
knitr::opts_chunk$set(echo = TRUE, results = "hold", tidy = TRUE)
```

<br />

```{r SimulationSetup}
##Setup variables provided by the instructions.
n <- 40 ##number of exponentials
lambda <- 0.2 ##universal lambda for all simulations
simNum <- 1000 ## number of simulations

set.seed(1234) ## set the seed to create reproducibility

##Simulation in matrix format
sim <- matrix(rexp(simNum * n, rate=lambda), simNum, n)
meanSim <- rowMeans(sim) ## Refer to Appendix
```


#### 1. Show the sample mean and compare it to the theoretical mean of the distribution.
***
As stated in the instruction the expected mean, mu of a exponential distribution is 1/lambda.We are going to compare this with the distribution mean.

<br />

```{r Theoretical mean}
muTheory <- 1/lambda
muTheory
```

<br />

```{r SampleMean}
muSim <-mean(meanSim)
muSim
```

<br />

The sample mean is `r muSim` and the theoretical mean is `r muTheory` which show that they are very close. This can be further seen with the histogram below.

<br />

```{r SampleMeanHistogram}
ggplot(data.frame(meanSim), aes(x = meanSim)) + 
  geom_histogram(position="identity", color = "black", fill="yellow", binwidth=0.4) + 
  labs(title = "Sample Mean Distribution Simulation", x = "Mean") +
  geom_vline(xintercept = muSim, size=1, colour="red") + 
  geom_vline(xintercept = muTheory, size=1, colour="green")
```

<br />

#### 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
***
To calculate variance of Exponential Distribution is Standard_Deviation^2/n

The theoretical variance is (1/lambda)^2/n

```{r VarianceThoery}
varTheory <- (1/lambda)^2/n
varTheory
```

The Sample variance is 
```{r VarianceSample}
varSample <- var(meanSim)
varSample
```

From the calculation above, we know that the sample distribution variance is `r varSample` and the theoretical variance is `r varTheory` 

<br />

#### 3. Show that the distribution is approximately normal.
***

The central limit theorem states that if you have a population with mean and standard deviation and take sufficiently large random samples from the population with replacement, then the distribution of the sample means will be approximately normally distributed.This means that we can use the normal probability model to quantify uncertainty when making inferences about a population mean based on the sample mean.

Below plot shows that the the sample mean distribution is approximately normal.



```{r SampleMeanNormalComparable}
ggplot(data.frame(meanSim), aes(x = meanSim,)) + 
  geom_histogram(position="identity", color = "black", fill="yellow", binwidth=0.4) + 
  geom_density(aes(y=0.4*..count..),colour="blue", size=1) +
  ##stat_function(fun = dnorm, colour = "green", args = list(mean = muTheory, sd=sqrt(varTheory))) +
  labs(title = "Sample Mean vs Normal Distribution", x = "Mean") + 
  geom_vline(xintercept = muSim, size=1, colour="red") + 
  geom_vline(xintercept = muTheory, size=1, colour="green")
```

<br />

#### 4. Appendix
***

First 6 lines of the mean of simulation
`r head(meanSim)`

Histogram of the Sample Mean Distribution


The platform specification used:

Spec    | Description
------- | -----------------------
OS      | Windows 10 Pro - 64 bit
CPU     | AMD Ryzen 5 - 3400G
RAM     | 16GB DDR4 3000MHz
Storage | 500GB SSD - M.2 NVMe (PCIe) 

